Conflict and Strategy in Collective Computation

### 2015-05-06 Brush_collectivecomputation_ms.tex
writing ms and supplemental and making pretty figures

\begin{document}
\begin{center}
\Large

\end{center}

\vspace{0pt}

\begin{center}
{\bf \LARGE{Conflict and Strategy in Collective Computation}}
\end{center}

\section{Abstract}
Groups of individuals making decisions in noisy environments perform collective computations in many biological systems, ranging in size from a brain made of billions of neurons to social groups made up of tens of monkeys. In each of these systems, pairs of individuals make decisions about their relative values. We use a leaky integrator model, a set of stochastic differential equations describing how decision variables change over time, to model the decision process. The decisions made between each pair form a directed network over the whole group. The degree of consensus in the group about each individual's value, as determined by the decision network, is a collective computation that is informative about the individuals' true values. We find that differences in the model when it is applied to neural and social systems do not affect the accuracy of decisions.  We find that conflicts of interest in the decision process affect individuals with low values the most and that conflicts of interest can incentivize these individuals to make more accurate decisions. We find that when pairwise decisions are accurate, local properties of the decision network are the most informative, but when accuracy decreases global properties are more informative. Finally, we find that the shape of the distribution of consensus scores is determined by the waiting costs of those decisions.

\section{Keywords }
collective computation, noisy learning, leaky integrator, decisions, social network, consensus, power, social structure

\section{Introduction} 
Many biological systems perform collective computation.  When presented with a set of moving dots, the brain, made up of billions of neurons, can decide whether the dots are moving left or right, even if the difference in the sizes of the groups moving either direction is not very large. A group of quorum sensing bacteria can sense its cell density and start to produce toxic, phosphorescent, or colorful molecules only when the density is sufficiently high. Flocks of birds and schools of fish can stay together and navigate with beautiful coordinated motion. In social groups of primates, the distribution of the degree of consensus about each monkeyÕs ability to use force is informative about the animalsÕ abilities and the animals use estimates of this emergent property to decide how to behave. In each of these examples, the individual members of the group make decisions based on noisy environmental information and a function of or algorithm acting on those individual decisions determines the groupÕs output. This two-step process allows the group to produce a successful output or behavior, despite computational limitations of its members and the conflicts of interest between them. While these examples come from quite different systems, the similarity in the phenomena suggest there may be common principles underlying collective computation in all of them.

 The sequential probability ratio test (SPRT) is the optimal way to decide between two alternatives in a noisy environment in the sense that for a given amount of time the SPRT is the most accurate algorithm and for a given error rate the SPRT reaches a decision most quickly.  The algorithm works by keeping track of the amount of evidence supporting one alternative and deciding in favor of that alternative once that amount is sufficiently high or deciding against it once that amount is sufficiently low. While this algorithm assumes the decider has perfect memory, in reality this is rarely the case and a simple way to model memory loss is to assume that the amount of evidence decays back toward 0 at teach point in time. This forgetful model of decision-making can be transformed into a leaky integrator model, a set of stochastic differential equations (SDEs) describing how the variables accumulating evidence for each alternative change over time. The random walk described by the SDEs is biased by the true state of the environment. This model is applicable to any system in which an individual with imperfect memory gathers evidence and chooses the alternative for which he has gathered more support.  The leaky integrator model, while a powerful for modeling noisy learning dynamics, has been applied rather phenomenologically, and we seek here to justify its use with a mechanistic derivation.

In this paper, we will study a group of individuals, each pair of which is trying to make a decision about their relative values. Each individual will have a decision variable indicating how much evidence it has accumulated and a threshold.  If the decision variable reaches the threshold, the individual will signal that a decision has been made. Both individuals involved in the decision value making a correct and fast decision. Since it takes time to gather enough evidence to make a correct decision, there is a tradeoff between accuracy and speed. Additionally, each individual might prefer that the decision be made in their favor, whether or not they truly have a higher value.  This introduces a conflict of interest. The thresholds are the strategies they use to negotiate accuracy, speed, and their personal preferences.

The leaky integrator model has been applied to neural decision-making, and it successfully explains and predict performance in a number of experiments  \citep{Eckhoff:2008uq, Brown:2005fk,Feng:2009kl,Bogacz:2006uq}. In this application, each of two neural populations responds to evidence for either left or right motion.  The SDEs describe how the firing rates of the two populations change over time and a decision is made once one the difference in firing rates is sufficiently high. Usually, the visual stimulus driving this decision process is a set of dots, some percentage of which are moving coherently either left or right and the rest of which are moving randomly.  The bias in the SDEs is in favor of the direction in which the coherent dots are moving.  However, this process could be used by neural populations deciding between any two possibilities, for example between up and down motion or between two colors, and a group of neural populations could be used to distinguish between or rank a whole set of possibilities.

While the SPRT has been applied to animal conflicts  \citep{Froment:2010fk}, to our knowledge the leaky integrator model has not been. In this application, each of two animals has a belief about its dominance with respect to the other which increases or decreases as the animal wins or loses fights against its opponent. The SDEs describe how these beliefs change over time. The source of evidence is the fight history between the two animals. Even if one animal is stronger and tends to win fights, the outcome of a fight is affected by the identity and presence of each animal's allies, motivation levels,  external environmental variables, and other stochastic factors, which introduce noise.   The bias in the SDEs is in favor of the animal that is more likely to win. Once an animals' opinion goes below a threshold, it decides to agree to be subordinate to the other. A correct decision is one in which the weaker individual makes this agreement.  In pigtailed macaques ({\it Macaca nemestrina}) and other species of macaques, there is a formal subordination signal that an animal can use to communicate its agreement to the subordinate role in their relationship \citep{Flack:2007kx, Flack:2006fk,Flack:2004oq, Waal:1985fk,Caldecott:1986uk,Preuschoft:2004ly}.  This signal marks the end of the decision process.

The leaky integrator model can thus be used to describe a decision being made between each pair of individuals in a group.  These decisions form a directed network, each edge of which reflects a decision about the relative values of the nodes it connects. In many systems, the degree of consensus in a group about each individuals' state is informative about the individuals'  true states and affects the individuals' functional behavior \cite{Brush:2013fk}. In previous work, we developed a number of network metrics to measure the degree of consensus in the group about each node and we found a few that produced scores that were predictive of individuals' functional behaviors \cite{Brush:2013fk,Flack:2006uq}. The output of one of these consensus measures being applied to a decision network would be a set of consensus scores, a collective computation which would be informative about the individuals' true states.

In the social case, this consensus is referred to as social power. In macaques, power is a useful variable since it reflects the animals' true abilities and the animals themselves do use estimates of their peers' and their own power to decide how to behave \cite{Flack:2004oq,Flack:2006fk,Flack:2006uq}.  The distribution of power scores in an empirical group of pigtailed macaques is heavy tailed and this allows the individuals in the tail to perform different functions in the social group than the rest of its members \cite{Flack:2005uq,Flack:2006fk,Flack:2006uq}.  Further, the shape of the power distribution varies across species of macaques \cite{Caldecott:1986uk,Flack:2004oq,Preuschoft:2004ly,Waal:1985fk,Thierry:2000ij}, and it is an open question whether this variation can be explained by differences in the underlying distributions of fighting abilities, by differences in the way animals in different species resolve their differences, or through other mechanisms. This kind of consensus may be important in neural systems, as well.  A number of neural populations may respond to or encode different environmental stimuli or variables.  The consensus in a group of neural populations about the value of a number of environmental stimuli could help the brain make decisions about the relative values of many possibilities.

There are thus three factors affecting the collective computation: the statistical properties of the environment, the strategies individuals use to make decisions about their relative values, and the method of measuring consensus from a network of decisions.  In this paper, first we derive the leaky integrator model from a simple mechanistic description of the accumulation of information, making two simplifying assumptions about the timescales on which probabilistic events occur \cite{Gillespie:2000fk}. We then use this model to answer the following questions. Do differences in the decision making model as it is applied to neural and social systems affect the accuracy of the decision process? What threshold strategies should individuals with conflicting interests use to make decisions in a group context? How do the individualsÕ threshold strategies affect the collective computation? What function of the decision network produces the most informative collective computation? How well does our model describe empirical social systems?

\section{Model} 
\subsection{Stochastic differential equations }    
%% make debt to Gillespie clearer

There are two possible types of environmental evidence: in the neural system, a dot can move either left or right and in the social system, a fight can be won by one animal or another.  For clarity in the following we will refer to these left or right dots, but the biological interpretation of the two types of inputs depends on the system. There are two decision variables, $X_1$ and $X_2$. In the absence of any input, the decision variables leak back towards $0$ with rate $\ell$.  (A table of all variables used in the text is given in Table \ref{variables}.)  If there is no input, then over a period length $\tau$ a decision variable decreases as $X(t+\tau)=(1-\ell\tau)X(t)$.

Each decision variable responds positively to one type of input and negatively to the other.  In the neural case, one neural population responds positively to left dots and negatively to right dots and the other neural population does the opposite.  In the social case, one animal's estimate of its own dominance increases when it wins fights and decreases when it loses fights, and the other animal's estimate does the opposite. Neither decision variable has perfect memory.  Specifically, $X_1$ increases by a boost $b$ when it sees a left dot and decrements by $-b$ when it sees a right dot, and conversely for $X_2$.     
To calculate the variables at time $t+\tau$, we  count how many times each of type of input occurred in the time since $t$ and add the changes resulting from these events to the background leaky estimate (as in \cite{Gillespie:2000fk}):
\begin{align*}
X_1(t+\tau)&=(1-\ell\tau)X_1(t)+b\times\# \text{ of left dots in }[t,t+\tau)-b\times\# \text{ of right dots in }[t,t+\tau)
\\ X_2(t+\tau)&=(1-\ell\tau)X_2(t)-b\times\# \text{ of left dots in }[t,t+\tau)+b\times\# \text{ of right dots in }[t,t+\tau). 
\end{align*}

@@

\subsection{Reaching a decision }
Each decision variable, $X_1$ and $X_2$, accumulates information about the stimulus.  The sign and size of the two variables indicate how much evidence there is for each possible decision outcome.  When a lot of evidence for the first possibility has been observed, $X_1$ will be large and positive, and similarly for the second possibility.  If the difference in the variables can be observed, the difference indicates the relative strengths of the evidence for either possibility. If $Y=X_1-X_2$, when $Y$ is large enough the system decides on $X_1$ and when $Y$ is low enough the system decides on $X_2$.  Specifically, there are two thresholds, $T_2$ and $T_1$ such that if $Y>T_1$ the decision is for $X_1$ and if $Y<-T_2$ the decision is for $X_2$.  However, it may be the case that the difference in the variables cannot be observed.  In this case, the decision depends on whether $X_1$ or $X_2$ hits a threshold first.  Again, there are two thresholds, $T_2$ and $T_1$, and if $X_1<-T_1$ the decision is for $X_1$ and if the $X_2<-T_2$ the decision is for $X_2$.

@@

\subsection{Utility of the decision }
The thresholds affect how long the decision takes and the probabilities with which each output is decided upon.  A good decision is one that reaches the correct output quickly, i.e. the expected time until a decision is reached (decision time, DT) and the probability that an incorrect decision is made (error rate, ER) are low.   However, it is impossible to minimize both simultaneously since waiting longer and accumulating more evidence makes the decision more accurate and conversely.  It may also be the case that each individual prefers one output over the other, regardless of which is ``correct''.  In the social case, an animal would prefer that its partner agreed to be subordinate, whether or not it is truly stronger.  In the neural case, different neural populations may receive different rewards if the decision reaches different outputs, regardless of which of the alternatives is ``true."  In these cases, a variable wants to maximize the probability that his preference is reached (probability of preference being chosen, $\text{PP}_i$).

To describe the tradeoffs between error rate, decision time, and preference, we quantify the utility of the decision process by introducing three weights, $w_1$, $w_2$, and $w_3$ such that $w_1+w_2+w_3=1$.  These weights describe how the three quantities are prioritized.  For individual $i$, we define
\begin{equation*}
U_{i}=w_1\text{ER}+w_2\text{DT}+w_3(1-\text{PP}_i).
\end{equation*}
The decision is better if $U_i$ is lower.  In the social system, $w_2$ can be interpreted as the cost of fighting since when $w_2$ is higher, the time spent fighting until a decision is reached is more costly.  Similarly, $w_3$ can be interpreted as the benefit from being the dominant animal in a pair. In both cases, $w_3$ represents how stubborn the individuals are about their preferences. If $w_3\gneq 0$ and the individuals prefer different outcomes, neither can simultaneously be satisfied since $\text{PP}_1=1-\text{PP}_2$.

As show in the Appendix, each of the three decision properties, ER, DT, and PP satisfies a partial differential equation that depends on the SDEs and the parameters of the model. There are analytical solutions to these equations when the system of SDEs is reduced to one dimension. However, we were not able to find an analytical solution for the full two-dimensional system and we therefore used numerical methods to solve the PDEs in this case.

\subsection{Optimal thresholds }
An individual-- either a neural population or a monkey-- can be engaged in either a single decision or in many decisions at once.  In the social system, for example, an animal wants to decide whether it is dominant or subordinate to every other member of its social group.  For any pair of individuals, each has a utility, $U_1$ and $U_2$, and consequently each has a Nash equilibrium threshold such that, if they choose those thresholds, neither has an incentive to choose another.  Similarly, in a group of individuals, each individual $i$ has a utility $U_{ij}$ from its decision process with individual $j$.  His total utility is given by the average of these, $\langle U_{ij}\rangle _j$. These lead to an optimized set of thresholds such that no individual has an incentive to choose another threshold.

Since an individual is trying to use the same algorithm to learn about every one else, he is constrained to have only one decision threshold for each of these decision processes.  In order to consistently assign the difficulties with which all pairs in a group make decisions, we assign each individual a value $a_i$ and the difficulty with which the pair $i,j$ makes a decision increases with the difference in abilities, i.e. $c_{ij}$ is an increasing function of $|a_i-a_j|$. In the social system, these values can be interpreted as the fighting abilities of the animals.The optimal decision threshold in this group context depends on the distribution of difficulties ($c_{ij}$) of the decision an individual will have to make and the thresholds of the other individuals.

 We use the following procedure to find optimal thresholds as a function of position in the group.  We repeatedly draw a set of fighting abilities from a uniform distribution; we find the optimal thresholds for each set, and then find the average optimal threshold as a function of an animal's order in the set of fighting abilities. To find the optimal thresholds for a given set of fighting abilities, we turn the difference in fighting abilities between each pair into the difficulty of the decision they have to make ($c$).  For an initial set of thresholds, we find $U_{ij}$ for each animal $i$ in each of its decisions and then average $\langle U_{ij}\rangle_j$ across all the animals with which it is engaged in a decision process.  We then allow each animal to choose a threshold that optimizes $\langle U_{ij}\rangle_j$ with respect to the rest of the group's thresholds and we iterate these choices until no animal chooses a new threshold.

\subsection{Collective computation }
We are interested in how the consensus about the value of each individual emerges out of the decision process.  (In the social system, an animal's social power depends on the consensus in the group about its fighting ability \cite{Brush:2013fk,Flack:2004oq,Flack:2006uq}.)  Given an interaction network, there are many ways to measure the consensus present in the group about each individual's value \cite{Brush:2013fk, Flack:2006uq}. These formalisms assign a consensus score to each node in the network (interpreted as a power score in the social system).  In the following we consider four  consensus formalisms.  The first and simplest is simply the number of nodes sending decisions to each animal (number of signalers).  The second is the number of decisions received by each node, which is a finer measure than the first.  The third is the entropy of the distribution of the number of decisions sent to each individual, which measures gives a coarse measurement of how uniformly all other animals in the group behave with respect to a focal individual.  The fourth is the eigenvector centrality of the signaling network, which measures how central each node is in the global structure of the network.  For more details see \cite{Brush:2013fk}.

To use these measures, we need to construct an interaction network between individuals in the group.  For each random set of fighting abilities, we let the animals optimize their thresholds, which give the probability that each pair will reach either decision and the expected time it will take.  We use these probabilities to generate a set of decisions between all pairs.  We use the convention that a decision is sent from $i$ to $j$ if in their decision $j$'s preference prevailed.  (In the social system, $i$ sends a subordination signal to $j$ to communicate his agreement to be subordinate.) To incorporate the time it takes for a decision to be reached and allow for multiple decisions to be sent, we say that $i$ sends $j$ a number of decisions equal to $1$ minus the expected decision time.  At each point in time $t$, the weight of the edge from $i$ to $j$ is $0$ if the expected decision time is greater than $t$ (i.e. a decision has not been made) and is given by the number of signals sent from $i$ to $j$ otherwise. Using many random sets of fighting abilities, for each consensus formalism and each point in time $t$, we can find the mutual information between the consensus scores from the network at $t$ and the true values. We can also find the average skewness of the set of consensus scores.

\section{Results}
\subsection{Decision making is as accurate without a third party arbiter. }
First we compare the decision making process in the full two-dimensional system and in a reduced one-dimensional system to see whether decisions are made more accurately or quickly in one or the other.  To make the systems directly comparable, we assume in both cases that the thresholds are symmetric, i.e. $T_1=T_2$.  We find that the two-dimensional decision is as accurate as the one-dimensional decision.  That is, for a given expected time to reach a decision, the two-dimensional decision will reach the correct decision with the same probability. Because they perform equally well and because the two-dimensional system is more appropriate for the social system (and perhaps the neural system as well), we use the two-dimensional system in the rest of our analyses. While the results come from an analysis of the two-dimensional system, they should also apply to the reduced system.
%Accuracy decreases as leak rates ($\ell$) become higher, but increases when the strength of the input ($c$) is higher.

\subsection{A conflict of interests incentivizes individuals in the bottom of the group to make more accurate decisions. }
The Nash equilibrium thresholds for a pair of individuals depend on whether the individuals are minimizing error rate, maximizing probability of their preference, or minimizing decision time.  If accuracy of the decision is the only thing that matters ($w_1=1$), the weaker individual will set his threshold as low as possible and the stronger animal will set his threshold as high as possible since if the decision is correct the weaker animal gives up first. When personal preference matters ($w_3>0)$ and there is a conflict of interest between the individuals, both will set their threshold high since there is no incentive for the individuals to stop accumulating evidence. As the importance of decision time increases (i.e. waiting costs increase), both individuals will lower their thresholds in order to reach the decision more quickly, whatever that decision may be (Figure \ref{nasheq}).  Notice that decision time and personal preference lead both animals to have the same thresholds whereas the individuals have different thresholds when accuracy is important. The accuracy with which a pair using Nash thresholds can reach a decision is highest when only error rate matters ($w_1=1$) and decreases as either decision time or personal preference become important (Figure \ref{nasheq}).

>>>>

\begin{document}

\title{Conflict and Strategy in Collective Computation}
\author{Eleanor Redstart Brush, David C. Krakauer, Jessica C. Flack}
\maketitle

\section*{Abstract}
Collective computations occur in many biological systems, ranging in size from a brain made of billions of neurons to social groups made up of tens of monkeys. In each of these systems, a group-level output emerges from the behavior of individuals making decisions in noisy environments. In particular, in the brain and in primate social groups, pairs of individuals must make decisions about their relative quality or ability. We use a leaky integrator model, a set of stochastic differential equations describing the dynamics of two decision variables, to model this decision process. The decisions made between each pair form a directed network over the whole group. The degree of consensus in the group about each individual's value, as determined by the decision network, is a collective computation that is informative about the individuals' true values. First, we show that the same model can be used to describe both neural and social systems. Next, we find that conflicts of interest in the decision process most strongly affect individuals at the bottom of the group and that conflicts of interest can incentivize these individuals to make more accurate decisions. Then we find that when pairwise decisions are accurate, local properties of the decision network are the most informative, but when accuracy decreases global properties are more informative. Finally, we find that the shape of the distribution of consensus scores is determined by the waiting costs of those decisions.

\section*{Keywords }
collective computation, noisy learning, leaky integrator, decisions, social network, consensus, power

\section*{Introduction} 
Many biological systems perform collective computation.  When presented with a set of moving dots, the brain, made up of billions of neurons, can decide whether more dots are moving left or right, even when the numerical superiority of one direction is very small. A group of quorum sensing bacteria can sense its cell density and start to produce toxic, phosphorescent, or colorful molecules once the density is sufficiently high. Flocks of birds and schools of fish can stay together and navigate with beautiful coordinated motion. In social groups of primates, the distribution of the degree of consensus about each monkey's ability to use force is informative about the animalsÕ abilities and the animals use estimates of this emergent property to decide how to behave. In each of these examples, the individual members of the group make decisions based on noisy environmental information and the collective computation emerges from those individual decisions. Through this two-step process the group produces a successful output or behavior, despite the computational limitations of its members and the conflicts of interest between them. While these examples come from quite different systems, the similarity in the phenomena suggest there may be common principles underlying collective computation in all of them.

%Both individuals involved in the decision value making a correct and fast decision. Since it takes time to gather enough evidence to make a correct decision, there is a tradeoff between accuracy and speed. Additionally, each individual might prefer that the decision be made in their favor, whether or not they truly have a higher ability.  This introduces a conflict of interest. Each individual will have a strategy to negotiate accuracy, speed, and their personal preferences.

In this paper, we study a group of individuals, each of which is trying to make a decision about its quality or ability relative to its peers. The sequential probability ratio test (SPRT) is the optimal way to decide between two alternatives in a noisy environment in the sense that for a given amount of time the SPRT is the most accurate algorithm and for a given error rate the SPRT reaches a decision most quickly.  The algorithm works by keeping track of the amount of evidence supporting each alternative and deciding in favor of one once the evidence for it is sufficiently high. This algorithm assumes the individuals making the decision have perfect memory, but in reality this is rarely the case.  A simple way to model memory loss is to assume that the amount of evidence decays back toward $0$ at teach point in time. This forgetful model of decision-making can be transformed into a leaky integrator model, a set of stochastic differential equations (SDEs) describing how the variables accumulating evidence for each alternative change over time. The random walk described by the SDEs is biased by the true state of the environment. This model is applicable to any system in which an individual with imperfect memory gathers evidence and chooses the alternative for which he has gathered more support.

The leaky integrator model has been applied to neural decision-making, and it successfully predicts performance in a number of experiments  \citep{Eckhoff:2008uq, Brown:2005fk,Feng:2009kl,Bogacz:2006uq}. In this system, each of two neural populations responds to evidence for either left or right motion.  The SDEs describe how the firing rates of the two populations change over time and a decision is made once one the difference in firing rates is sufficiently high. Usually, the visual stimulus driving this decision process is a set of dots, some percentage of which are moving coherently either left or right and the rest of which are moving randomly.  The bias in the SDEs is in favor of the direction in which the coherent dots are moving.  The model could also describe neural populations deciding between other possibilities, for example between up and down motion or between two colors. In order to distinguish between or rank a whole set of possibilities, the brain could rely on a group of neural populations, each responding to one of the options.

While the SPRT has been applied to animal conflicts  \citep{Froment:2010fk}, to our knowledge the leaky integrator model has not been. In this application, each of two animals has an opinion about its dominance with respect to the other which goes up or down as the animal wins or loses fights against its opponent. Even if one animal is stronger and tends to win fights, the outcome of a fight is affected by the identity and presence of each animal's allies, motivation levels,  external environmental variables, and other stochastic factors, which introduce noise that is captured by the SDEs.   The bias in the model is in favor of the animal that is more likely to win. Once an animals' opinion goes below a threshold, it decides to agree to be subordinate to the other. A correct decision is one in which the weaker individual makes this agreement.  In pigtailed macaques ({\it Macaca nemestrina}), there is a formal subordination signal that an animal can use to communicate its agreement to the subordinate role in their relationship \citep{Flack:2007kx, Flack:2006fk,Flack:2004oq, Waal:1985fk,Caldecott:1986uk,Preuschoft:2004ly}.  This signal marks the end of the decision process.

In both systems, the decisions between pairs of individuals form a directed network, each edge of which reflects a decision about the relative values of the nodes it connects. In many systems, the degree of consensus in a group about each individual's value is informative about the individuals'  true values and affects the individuals' functional behaviors \cite{Brush:2013fk}. In previous work, we developed a number of network metrics to measure the degree of consensus in the group about each node and we found a few that produced scores that were predictive of individuals' functions in the group \cite{Brush:2013fk,Flack:2006uq}. The consensus scores that result from applying one of these consensus measures to the decision network is the collective computation that we are interested in. There are three factors affecting this collective computation: the statistical properties of the environment, the strategies individuals use to make decisions about their relative values, and the method of measuring consensus from a network of decisions.

In the social system, this consensus is referred to as social power. In macaques, power is a useful variable since it reflects the animals' true fighting abilities and the animals themselves use estimates of their peers' and their own power to decide how to behave \cite{Flack:2004oq,Flack:2006fk,Flack:2006uq}.  The distribution of power scores in a group of pigtailed macaques is heavy tailed and this allows the individuals in the tail to perform different functions  than the rest of the animals \cite{Flack:2005uq,Flack:2006fk,Flack:2006uq}.  The shape of the power distribution varies across species of macaques \cite{Caldecott:1986uk,Flack:2004oq,Preuschoft:2004ly,Waal:1985fk,Thierry:2000ij}, and it is an open question whether this variation can be explained by differences in the underlying distributions of fighting abilities, by differences in the way animals in different species resolve their differences, or through other mechanisms. This kind of consensus may be important in neural systems, as well.  A number of neural populations may respond to or encode different environmental stimuli or variables.  The consensus in a group of neural populations about the value of a number of environmental stimuli could help the brain make decisions about the relative values of many possibilities.

In this paper, we derive the leaky integrator model from a simple mechanistic description of the accumulation of information, making two simplifying assumptions about the timescales on which probabilistic events occur \cite{Gillespie:2000fk}. While the model has been successful used to describe neural learning dynamics, its application has been rather phenomenological and this mechanistic derivation justifies its application in any system where the assumptions are not unreasonable. We then use the leaky integrator model to answer the following questions. Do differences in the decision making model as it is applied to neural and social systems affect the accuracy of the decision process? What threshold strategies should individuals with conflicting interests use to make decisions in a group context? How do the individualsÕ threshold strategies affect the collective computation? What function of the decision network produces the most informative collective computation? How well does our model describe empirical social systems?

\section*{Model} 
\subsection*{Stochastic differential equations }    
%% make debt to Gillespie clearer
\ In a decision between two alternatives, there are two decision variables, $X_1$ and $X_2$, each indicating the evidence accumulated for one of the options.  In the absence of any input, the decision variables leak back towards $0$ with rate $\ell$.  (A table of all variables used in the text is given in Table \ref{variables}.)  If there is no input, then over a period length $\tau$ a decision variable decreases as $X(t+\tau)=(1-\ell\tau)X(t)$. If there is input, each decision variable will incorporate the new evidence by responding positively to one type of input and negatively to the other.  Specifically, $X_1$ increases by a boost $b$ when it sees a left dot and decrements by $-b$ when it sees a right dot, and conversely for $X_2$.  In the neural case, one neural population responds positively to left dots and negatively to right dots and the other neural population does the opposite.  In the social case, one animal's estimate of its own dominance increases when it wins fights and decreases when it loses fights, and the other animal's estimate does the opposite.  For clarity, in the following we will refer to the inputs as left or right dots, but they can be interpreted as wins or losses in the social system. To calculate the variables at time $t+\tau$, we  count how many times each of type of input occurred in the time since $t$ and add the changes resulting from these events to the background leaky estimate (as in \cite{Gillespie:2000fk}):
\begin{align*}
X_1(t+\tau)&=(1-\ell\tau)X_1(t)+b\times\# \text{ of left dots in }[t,t+\tau)-b\times\# \text{ of right dots in }[t,t+\tau)
\\ X_2(t+\tau)&=(1-\ell\tau)X_2(t)-b\times\# \text{ of left dots in }[t,t+\tau)+b\times\# \text{ of right dots in }[t,t+\tau). 
\end{align*}

@@

\subsection*{Reaching a decision }\ There are two ways to use the decision variables $X_1$ and $X_2$ to actually make a decision.  If the difference in the variables can be observed, the difference indicates the relative strengths of the evidence for either possibility. If $Y=X_1-X_2$, the system should decides on $X_1$ when $Y$ is large and positive on $X_2$ when $Y$ is large and negative.  Specifically, there are two thresholds, $T_1$ and $T_2$ such that if $Y>T_1$ the decision is for $X_1$ and if $Y<-T_2$ the decision is for $X_2$.  However, it may be the case that the difference in the variables cannot be observed.  In this case, the decision depends on whether $X_1$ or $X_2$ hits a threshold first.  Again, there are two thresholds, $T_1$ and $T_2$, and if $X_2<-T_2$ the decision is for $X_1$ and if the $X_1<-T_1$ the decision is for $X_2$.

@@

\subsection*{Utility of the decision }
\ The thresholds affect how long the decision takes and the probabilities with which each output is decided upon.  A good decision is one that reaches the correct output quickly, i.e.\ the expected time until a decision is reached (decision time, DT) and the probability that an incorrect decision is made (error rate, ER) are low.   However, it is impossible to minimize both simultaneously since waiting longer and accumulating more evidence makes the decision more accurate and conversely.  It may also be the case that each individual prefers one output over the other, regardless of which is ``correct''.  In the social case, each animal would prefer that its partner agree to be subordinate, whether or not it is truly stronger.  In the neural case, different neural populations may receive different rewards if the decision reaches different outcomes, regardless of which of the alternatives is ``correct."  Therefore, an individual may want to maximize the probability that its preferred outcome is reached (probability of preference being chosen, $\text{PP}_i$).

To describe the tradeoffs between error rate, decision time, and preference, we quantify the utility of the decision process by introducing three weights, $w_1$, $w_2$, and $w_3$ such that $w_1+w_2+w_3=1$.  These weights describe how the three quantities are prioritized.  For individual $i$, we define
\begin{equation*}
U_{i}=w_1\text{ER}+w_2\text{DT}+w_3(1-\text{PP}_i).
\end{equation*}
The decision is better if $U_i$ is lower.  In the social system, $w_2$ can be interpreted as the cost of fighting since when $w_2$ is higher, the time spent fighting until a decision is reached is more costly.  In the neural system, $w_2$ depends on whatever costs the brain or the whole animal incurs by waiting for a decision. The personal preference weight $w_3$ indicates the strength of the conflict of interests between individuals since it is impossible for them both to maximize $\text{PP}_i$, given that $\text{PP}_1=1-\text{PP}_2$. In the social system, $w_3$ can be interpreted as the benefit from being the dominant animal in a pair. In both systems, $w_3$ represents how stubborn the individuals are about their preferences.

Each of the three decision properties, ER, DT, and PP satisfies a partial differential equation that depends on the SDEs and the parameters of the model. There are analytical solutions to these equations when the system of SDEs is reduced to one dimension. However, we were not able to find an analytical solution for the full two-dimensional system and we therefore used numerical methods to solve the PDEs in this case.

\subsection*{Nash equilibrium thresholds }
\ In a group of individuals, each with a value $a_i$, the difficulty with which the pair $i,j$ makes a decision increases with the difference in value, i.e. $c_{ij}$ is an increasing function of $|a_i-a_j|$. In the social system, the value $a_i$ reflects an animal's fighting ability. In the neural system, the value $a_i$ is high if the property (e.g.\ left movement or red color) a neural population responds to is abundant in the visual stimulus. We assume that each individual has the same decision threshold for all the decisions processes with each of its peers and, given those thresholds, each individual $i$ has a utility $U_{ij}$ from its decision process with individual $j$ and a total utility given by the average of these, $\langle U_{ij}\rangle _j$. For each set of values $\{a_1,\dots,a_N\}$, we find the Nash equilibrium thresholds such that no individual has an incentive to choose another threshold to improve its utility.  Since the Nash thresholds depend on the values $\{a_1,\dots,a_N\}$,  we repeatedly draw a set of values from a uniform distribution and find the Nash thresholds for each set. Then we find the average Nash threshold as a function of an animal's rank in the set of fighting abilities.

\subsection*{Collective computation }
\ For each set of values $\{a_1,\dots,a_N\}$ and Nash thresholds $\{T_1,\dots,T_N\}$, we find the probability that each pair will reach either outcome and the expected time it will take.  We use these probabilities to generate a set of decisions between all pairs.  We use the convention that a decision is sent from $i$ to $j$ if $j$'s preference prevailed.  To incorporate the time it takes for a decision to be reached and allow for multiple decisions to be made, we say that $i$ sends $j$ a number of decisions equal to $1$ minus the expected decision time.  At each point in time $t$, the weight of the edge from $i$ to $j$ is $0$ if the expected decision time is greater than $t$ (i.e. a decision has not been made) and is given by the number of signals sent from $i$ to $j$ otherwise.

Given a decision network, there are many ways to measure the consensus present in the group about each individual's value \cite{Brush:2013fk, Flack:2006uq}. These formalisms assign a consensus score to each node in the network.  In the following we consider four  consensus formalisms.  The first and simplest is simply the unweighted in-degree of a node, i.e. the number of individuals deciding in favor of each individual.  The second is the weighted in-degree of a node, i.e. the number of decisions made in favor of each node, which is a finer measure than the first.  The third is the entropy of the distribution of the number of decisions made in favor of each individual, which gives a coarse measurement of how uniformly all other individuals in the group behave with respect to a focal individual.  The fourth is the eigenvector centrality of the decision network, which measures how central each node is in the global structure of the network.  For more details see \cite{Brush:2013fk}. Using many random sets of fighting abilities, we find the mutual information between the consensus scores from the network at each time $t$ and the true values, for each consensus formalism. We also find the average skewness of the set of consensus scores.

\section*{Results}
\subsection*{Decision making is as accurate without a third party arbiter. }
\ First we compare the decision making process in the full two-dimensional system and in a reduced one-dimensional system to see whether decisions are made more accurately or quickly in one or the other.  To make the systems directly comparable, we assume in both cases that the thresholds are symmetric, i.e. $T_1=T_2$.  We find that the two-dimensional decision is as accurate as the one-dimensional decision.  That is, for a given expected time to reach a decision, the two-dimensional decision will reach the correct decision with the same probability. Because they perform equally well and because the two-dimensional system is more appropriate for the social system (and perhaps the neural system as well), we use the two-dimensional system in the rest of our analyses. While the results come from an analysis of the two-dimensional system, they should also apply to the reduced system.
%Accuracy decreases as leak rates ($\ell$) become higher, but increases when the strength of the input ($c$) is higher.

\subsection*{A conflict of interests incentivizes individuals in the bottom of the group to make more accurate decisions. }
\ We start with a pair of individuals, i.e. a group of size two, to build intuition for how the optimization weights affect the Nash equilibrium thresholds.  If accuracy of the decision is the only thing that matters ($w_1=1$), the weaker individual will set his threshold as low as possible and the stronger animal will set his threshold as high as possible since if the decision is correct the weaker animal gives up first. When personal preference matters ($w_3>0)$ and there is a conflict of interest between the individuals, both will set their threshold high since there is no incentive for the individuals to stop accumulating evidence. As the importance of decision time increases (i.e. waiting costs increase), both individuals will lower their thresholds in order to reach the decision more quickly, whatever that decision may be (Figure \ref{nasheq}).  Notice that decision time and personal preference lead both animals to have the same thresholds whereas the individuals have different thresholds when accuracy is important. The accuracy with which a pair using Nash thresholds can reach a decision is highest when only error rate matters ($w_1=1$) and decreases as either decision time or personal preference become important (Figure \ref{nasheq}).

In a group with more than two individuals, an individual's position in the group affects how it makes the tradeoff between error rate, preference and decision time.  As in a pair, prioritizing accuracy tends to lead animal's at the top of the group to set high thresholds and those at the bottom to set low thresholds. Prioritizing personal preferences, on the other hand, tends to lead all individuals to set high thresholds (Figure \ref{groupeq}). The difference in an individual's threshold between when accuracy is important and when personal preferences are important is largest for the individuals at the bottom of the group. Individuals towards the top will set high thresholds in either case. Individuals toward the bottom, however, will set lower thresholds when they prioritize accuracy then when they prioritize personal preferences. This is because choosing a higher threshold would increase waiting time in decisions with respect to all individuals and decrease the accuracy of its decisions with respect to higher animals, and it would only slightly increase the accuracy of its decisions with respect to other low individuals. In other words, individuals at the bottom of the group sacrifice some accuracy in their decisions with respect to each other in order to decrease their overall decision times.
